Detecting Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

Our initial results using the Pegasos SVM solver included in ScalaNLP were disappointing. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%. Table \{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
"2012," "wish," and "you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of January.

@@

The ham message collection methodology used in this project may have been flawed, leading questionable 
performance for several classifiers. The 2 million ham messages were sent over a span of 10 days, from 
December 23, 2011 to January 1, 2012. While we chose the corpus' start date as exactly four months before 
extracting the data, the coincidental timing with a global holiday seems to have introduced unusual 
skew in the bag-of-words features. In contrast, the majority of messages in the spam corpus span a period of 
7 months, with only a slight linear increase in volume over the interval.  Consequently, as seen in 
the Linear Regression and SVM feature weight tables, words typically associated with the New Year gained 
undue wieght during classification.  We plan to address this issue by creating a new ham message corpus 
through sampling that matches the temporal distribution of spam messages more closely.

>>>>

Our initial results using the Pegasos SVM solver included in ScalaNLP were disappointing. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
"2012," "wish," and "you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of January.

@@

The ham message collection methodology used in this project may have been flawed, leading to questionable 
performance for several classifiers. The 2 million ham messages were sent over a span of 10 days, from 
December 23, 2011 to January 1, 2012. While we chose the corpus' start date as exactly four months before 
extracting the data, the coincidental timing with a global holiday seems to have introduced unusual 
skew in the bag-of-words features. In contrast, the majority of messages in the spam corpus span a period of 
7 months, with only a slight linear increase in volume over the interval.  Consequently, as seen in 
the Linear Regression and SVM feature weight tables, words typically associated with the New Year gained 
undue wieght during classification.  We plan to address this issue by creating a new ham message corpus 
through sampling that matches the temporal distribution of spam messages more closely.
